---
title: "CH4 initial"
author: 
Zhang Zhenglin
output:
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: sentence
---

# Necessary libraries

```{r message=FALSE,warning=FALSE}
library(knitr)
library(ggplot2)
theme_set(theme_bw())
library(emmeans)
library(multcomp)
library(PLS205)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(Rmisc) 
library(dplyr) #https://r4ds.had.co.nz/ (Chapter 3, Chapter 5, look at filter and select)
# https://bookdown.org/ansellbr/WEHI_tidyR_course_book/
library(stringr) 
library(data.table)
library(GGally)
library(formatR)
library(readxl)
library(FluxCalR)
library(tidyverse)
library(fuzzyjoin)
library(purrr)
library(data.table)
library(broom)
library(lubridate)
library(readxl)
library(openxlsx)
library(ggrepel)
library(gridExtra)
library(ggpubr)
library(cluster)
```

# Load data

```{r}
master <- read_excel("../Data/Initial_Database.xlsx", sheet = 1)
str(master)
```

# Make sure all columns are in the correct variable class

```{r}
master$Number <- as.factor(master$Number)
```


# Convert all observations to common currency

# Linear interpolation function

Number column here is our biological replicate. One set of time resolved flux data for one cumulative flux value. 


```{r}
library(dplyr)

interpolate_emissions <- function(data,
                                  group_cols = c("Number", "Year", "Location", "Variety"),
                                  day_col = "Days_After_Flooding",
                                  emission_col = "Emission_g_CH4_ha_day") {

  data %>%
    arrange(across(all_of(c(group_cols, day_col)))) %>%
    group_by(across(all_of(group_cols))) %>%
    group_modify(~ {

      df_group <- .x

      # Remove NA emission rows for interpolation check
      valid_rows <- df_group %>%
        filter(!is.na(.data[[emission_col]]))

      # If fewer than 2 valid points → cannot interpolate
      if (nrow(valid_rows) < 2) {
        return(df_group)
      }

      # Full daily sequence
      full_days <- seq(min(valid_rows[[day_col]]),
                       max(valid_rows[[day_col]]),
                       by = 1)

      # Linear interpolation
      interp_vals <- approx(
        x = valid_rows[[day_col]],
        y = valid_rows[[emission_col]],
        xout = full_days,
        method = "linear"
      )

      # Create template from first row
      template <- df_group[1, ]
      template <- template[rep(1, length(full_days)), ]

      template[[day_col]] <- interp_vals$x
      template[[emission_col]] <- interp_vals$y

      template
    }) %>%
    ungroup()
}
```

# Linear interpolation

```{r}
flux_interpolated <- interpolate_emissions(master) %>% mutate(Day_Int = as.integer(round(Days_After_Flooding)))

```



# Interpolation visualisation

## check to make sure code is functioning well

```{r}
ggplot(flux_interpolated%>%filter(Number=="1"), aes(y=Emission_g_CH4_ha_day, x=Days_After_Flooding, color = Number)) + geom_point()

ggplot(master%>%filter(Number=="1"), aes(y=Emission_g_CH4_ha_day, x=Days_After_Flooding, color = Number)) + geom_point()

```


```{r}
ggplot(flux_interpolated, aes(y=Emission_g_CH4_ha_day, x=Days_After_Flooding, color = Number)) + geom_point()

ggplot(master, aes(y=Emission_g_CH4_ha_day, x=Days_After_Flooding, color = Number)) + geom_point()

```

# Correlation function (look at corr of Emission_g_CH4_ha_day with cumulative)

```{r}
cor_emission_cumulative_by_day <- function(df,
                                           emission_col = "Emission_g_CH4_ha_day",
                                           cumulative_col = "Cumulative_seaonal_kg_CH4_ha",
                                           day_col = "Days_After_Flooding") {
  
  df %>%
    mutate(
      Day_Int = as.integer(round(.data[[day_col]])),
      Emission_num = as.numeric(.data[[emission_col]]),
      Cumulative_num = as.numeric(.data[[cumulative_col]])
    ) %>%
    group_by(Day_Int) %>%
    summarise(
      n = sum(!is.na(Emission_num) & !is.na(Cumulative_num)),
      
      correlation = if (n >= 3) {
        cor(Emission_num,
            Cumulative_num,
            use = "complete.obs",
            method = "pearson")
      } else NA_real_,
      
      p_value = if (n >= 3) {
        cor.test(Emission_num,
                 Cumulative_num,
                 method = "pearson")$p.value
      } else NA_real_,
      
      .groups = "drop"
    )
}
```

# Try function 

```{r}
cor_results <- cor_emission_cumulative_by_day(flux_interpolated)

ggplot(cor_results, aes(y=correlation, x=Day_Int)) + geom_point()
```

# Try R2 and regression

## Function
```{r}
find_best_predictive_day <- function(df,
                                     emission_col = "Emission_g_CH4_ha_day",
                                     cumulative_col = "Cumulative_seaonal_kg_CH4_ha",
                                     day_col = "Days_After_Flooding") {
  
  results <- df %>%
    mutate(
      Day_Int = as.integer(round(.data[[day_col]])),
      Emission_num = as.numeric(.data[[emission_col]]),
      Cumulative_num = as.numeric(.data[[cumulative_col]])
    ) %>%
    group_by(Day_Int) %>%
    summarise(
      n = sum(!is.na(Emission_num) & !is.na(Cumulative_num)),
      
      model = if (n >= 3) list(lm(Cumulative_num ~ Emission_num)) else list(NULL),
      
      R2 = if (!is.null(model[[1]])) summary(model[[1]])$r.squared else NA_real_,
      
      slope = if (!is.null(model[[1]])) coef(model[[1]])[2] else NA_real_,
      
      p_value = if (!is.null(model[[1]])) 
        summary(model[[1]])$coefficients[2,4] else NA_real_,
      
      .groups = "drop"
    ) %>%
    select(-model)
  
  best_day <- results %>%
    filter(R2 == max(R2, na.rm = TRUE))
  
  list(
    all_days = results,
    best_day = best_day
  )
}
```

## Run dataset 

```{r}
output <- find_best_predictive_day(flux_interpolated)

all_days <- output$all_days      # full summary
output$best_day      # day with highest R²

ggplot(all_days, aes(y=R2, x=Day_Int)) + geom_point() + geom_hline(yintercept=0.95) + geom_hline(yintercept=0.905)

```
# Use Day 62 to check 

```{r}
ggplot(data=flux_interpolated%>%filter(Day_Int==62), aes(x=Emission_g_CH4_ha_day, y=Cumulative_seaonal_kg_CH4_ha))+
  geom_point()+
  geom_smooth(method = lm)+
  stat_regline_equation(aes(x=Emission_g_CH4_ha_day, y=Cumulative_seaonal_kg_CH4_ha, 
    label =  paste(..rr.label..)),
    show.legend = FALSE, 
    label.x = 2000,
    label.y = 100)+
  stat_regline_equation(aes(x=Emission_g_CH4_ha_day, y=Cumulative_seaonal_kg_CH4_ha, 
    label = paste(..eq.label..)),
    show.legend = FALSE, 
    label.x = 2000, 
    label.y = 150)
```
# Let's use day 62 to predict seasonal cumulative emissions.




# What if we just drew a triangle from day 62?

Essentially, if we just took the flux from day 62, linear interpolated between first and last days using the known growth period, and then summed daily emissions, how well will this predict seasonaly cumulative emissions?

Assume 130 days here for fun as a thought experiment

```{r}
interpolate_day62_triangle <- function(df,
                                       day_col = "Days_After_Flooding",
                                       emission_col = "Emission_g_CH4_ha_day",
                                       group_col = "Number") {
  
  df %>%
    mutate(
      Day_Int = as.integer(round(.data[[day_col]])),
      Emission_num = as.numeric(.data[[emission_col]])
    ) %>%
    group_by(.data[[group_col]]) %>%
    summarise({
      
      # Extract ONLY day 62 flux
      y62 <- Emission_num[Day_Int == 62][1]
      
      if (is.na(y62)) {
        tibble(total_emission_0_130 = NA_real_)
      } else {
        
        days <- 0:130
        
        interp_values <- ifelse(
          days <= 62,
          y62 * (days / 62),
          y62 * (130 - days) / (130 - 62)
        )
        
        tibble(
          total_emission_0_130 = sum(interp_values)
        )
      }
      
    }, .groups = "drop")
}

```

```{r}
day62_predict <- interpolate_day62_triangle(flux_interpolated) %>% mutate(total_emission_0_130_kg_ha = total_emission_0_130/1000)
```

## Join data

```{r}
master_summary <- master %>%
  group_by(Number) %>%
  summarise(Cumulative_seaonal_kg_CH4_ha = mean(Cumulative_seaonal_kg_CH4_ha, na.rm = TRUE),
            .groups = "drop")

master_joined <-day62_predict %>%
  left_join(master_summary, by = "Number")

master_joined$diff_triangle <- master_joined$total_emission_0_130_kg_ha-master_joined$Cumulative_seaonal_kg_CH4_ha
master_joined$diff_triangle_percent <- 100*abs(master_joined$diff_triangle/master_joined$Cumulative_seaonal_kg_CH4_ha)

mean(master_joined$diff_triangle_percent)

# generally, we see that this approach works, but it tends to underpredict emissions. But honestly, its a pretty damn good way to do it. 

ggplot(master_joined, aes(y=Cumulative_seaonal_kg_CH4_ha, x=total_emission_0_130_kg_ha)) + geom_point()+
geom_smooth(method = lm)+
  stat_regline_equation(aes(x=total_emission_0_130_kg_ha, y=Cumulative_seaonal_kg_CH4_ha, 
    label =  paste(..rr.label..)),
    show.legend = FALSE, 
    label.x = 200,
    label.y = 100)+
  stat_regline_equation(aes(x=total_emission_0_130_kg_ha, y=Cumulative_seaonal_kg_CH4_ha, 
    label = paste(..eq.label..)),
    show.legend = FALSE, 
    label.x = 200, 
    label.y = 150)+
    geom_abline(intercept = 0, slope = 1, linetype = "dashed")

```
# What is we used the linear regrssion of day 62 emission against seasonal cumulative to get a cumulative value?

```{r}

lm_day62 <- lm(Cumulative_seaonal_kg_CH4_ha~Emission_g_CH4_ha_day, flux_interpolated%>%filter(Day_Int==62))


master_joined$predict_linear_reg_ppm <- predict(lm_day62, flux_interpolated%>%filter(Day_Int==62))

```

```{r}
ggplot(master_joined, aes(y=Cumulative_seaonal_kg_CH4_ha, x=predict_linear_reg_ppm)) + geom_point()+
#geom_smooth(method = lm)+
  stat_regline_equation(aes(x=predict_linear_reg_ppm, y=Cumulative_seaonal_kg_CH4_ha, 
    label =  paste(..rr.label..)),
    show.legend = FALSE, 
    label.x = 200,
    label.y = 100)+
 # stat_regline_equation(aes(x=predict_linear_reg_ppm, y=Cumulative_seaonal_kg_CH4_ha, 
 #   label = paste(..eq.label..)),
 #   show.legend = FALSE, 
 #   label.x = 200, 
 #   label.y = 150)+
    geom_abline(intercept = 0, slope = 1, linetype = "dashed")

master_joined$diff_linear_reg <- master_joined$predict_linear_reg_ppm-master_joined$Cumulative_seaonal_kg_CH4_ha
master_joined$diff_linear_reg_percent_diff <- 100*abs(master_joined$diff_linear_reg/master_joined$Cumulative_seaonal_kg_CH4_ha)


mean(master_joined$diff_linear_reg_percent_diff)
```

