---
title: "CH4 initial"
author: 
Zhang Zhenglin
output:
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: sentence
---

# Necessary libraries

```{r message=FALSE,warning=FALSE}
library(knitr)
library(ggplot2)
theme_set(theme_bw())
library(emmeans)
library(multcomp)
library(PLS205)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(Rmisc) 
library(dplyr) #https://r4ds.had.co.nz/ (Chapter 3, Chapter 5, look at filter and select)
# https://bookdown.org/ansellbr/WEHI_tidyR_course_book/
library(stringr) 
library(data.table)
library(GGally)
library(formatR)
library(readxl)
library(FluxCalR)
library(tidyverse)
library(fuzzyjoin)
library(purrr)
library(data.table)
library(broom)
library(lubridate)
library(readxl)
library(openxlsx)
library(ggrepel)
library(gridExtra)
library(zoo)
```

# Load data

```{r}
master <- read_excel("../Data/Initial_Database.xlsx", sheet = 1)
str(master)
```

# Make sure all columns are in the correct variable class

```{r}
master$Number <- as.factor(master$Number)
```


# Convert all observations to common currency

# Linear interpolation function

Number column here is our biological replicate. One set of time resolved flux data for one cumulative flux value. 


```{r}
library(dplyr)

interpolate_emissions <- function(data,
                                  group_cols = c("Number", "Year", "Location", "Variety"),
                                  day_col = "Days_After_Flooding",
                                  emission_col = "Emission_g_CH4_ha_day") {

  data %>%
    arrange(across(all_of(c(group_cols, day_col)))) %>%
    group_by(across(all_of(group_cols))) %>%
    group_modify(~ {

      df_group <- .x

      # Remove NA emission rows for interpolation check
      valid_rows <- df_group %>%
        filter(!is.na(.data[[emission_col]]))

      # If fewer than 2 valid points → cannot interpolate
      if (nrow(valid_rows) < 2) {
        return(df_group)
      }

      # Full daily sequence
      full_days <- seq(min(valid_rows[[day_col]]),
                       max(valid_rows[[day_col]]),
                       by = 1)

      # Linear interpolation
      interp_vals <- approx(
        x = valid_rows[[day_col]],
        y = valid_rows[[emission_col]],
        xout = full_days,
        method = "linear"
      )

      # Create template from first row
      template <- df_group[1, ]
      template <- template[rep(1, length(full_days)), ]

      template[[day_col]] <- interp_vals$x
      template[[emission_col]] <- interp_vals$y

      template
    }) %>%
    ungroup()
}
```

# Linear interpolation

```{r}
flux_interpolated <- interpolate_emissions(master)
```



# Interpolation visualisation

```{r}
ggplot(flux_interpolated, aes(y=Emission_g_CH4_ha_day, x=Days_After_Flooding, color = Number)) + geom_point()
```
# Correlation function (look at corr of Emission_g_CH4_ha_day with cumulative)

```{r}
cor_emission_cumulative_by_day <- function(df,
                                           emission_col = "Emission_g_CH4_ha_day",
                                           cumulative_col = "Cumulative_seaonal_kg_CH4_ha",
                                           day_col = "Days_After_Flooding") {
  
  df %>%
    mutate(
      Day_Int = as.integer(round(.data[[day_col]])),
      Emission_num = as.numeric(.data[[emission_col]]),
      Cumulative_num = as.numeric(.data[[cumulative_col]])
    ) %>%
    group_by(Day_Int) %>%
    summarise(
      n = sum(!is.na(Emission_num) & !is.na(Cumulative_num)),
      
      correlation = if (n >= 3) {
        cor(Emission_num,
            Cumulative_num,
            use = "complete.obs",
            method = "pearson")
      } else NA_real_,
      
      p_value = if (n >= 3) {
        cor.test(Emission_num,
                 Cumulative_num,
                 method = "pearson")$p.value
      } else NA_real_,
      
      .groups = "drop"
    )
}
```

# Try function 

```{r}
cor_results <- cor_emission_cumulative_by_day(flux_interpolated)

ggplot(cor_results, aes(y=correlation, x=Day_Int)) + geom_point()

```



